Data Management for LSST Special Programs

### 2023-11-08 sec_sci.tex
add observing mode term

\subsection{Region and Program Labels}\label{ssec:sci_labels}

\textbf{All visits should have a region label \textit{and} a program 
label, and these labels should be propagated to processed images and catalog data.}

\textbf{Region Labels -- } based on Right Acension and Declination.
Region labels should include, e.g., WFD low-dust, Galactic Plane, Galactic 
Bulge, SCP, NES, or Virgo; DDF fieldname; mini-survey region name.
Region labels should be propagated to processed images (visit, 
difference, and deep coadds), catalog sources, and catalog static-sky objects.

\textbf{Program Labels -- } 
Based on survey cadeance or strategy.
Program labels might include, e.g., WFD, DDF, TOO, mini-survey name, or other 
options like engineering, commissioning, and director's discretionary.
Program labels should be propagated to single-visit processed and 
difference images and catalog sources.
It would not make sense to propagate program labels to deep coadds or 
object tables, as these data products could be a mix of programs.

@@

\item To enable provenance when Special Programs data in included in regular
processing and the WFD program data products (e.g., if used to improve the
all-sky coadd).

@@

\textbf{Necessary -- } When the primary science goal for a Special Program 
cannot be met by including the data in regular processing (e.g., Prompt processing),
or where doing so would compromise the WFD data products (e.g., introduce non-uniformity).

@@

The main motivation for this recommendation is that all time-domain and 
moving-object science goals (two of the four science pillars for the LSST) are
enhanced by any and all additional observations, even if they are not
optimized for the science (as the WFD program's cadence is being optimized).

@@

Users will need to be able to query for data that restricts by sky region and 
program label. 
This can be accomplished by including those labels in all image and catalog 
metadata as described in Section~\ref{ssec:sci_labels}, as the TAP service 
and butler already provide the mechanism for user-specified queries.

Users will need to be able to discover Special Programs data products when 
browsing data.
This could be accomplished with, e.g., a toggle to overlay region boundaries
for Special Programs when browsing all-sky coadds built from WFD program data
in the Portal Aspect.

>>>>

\subsection{Region and Observing Mode Labels}\label{ssec:sci_labels}

\textbf{All visits should have a region label \textit{and} an observing mode 
label, and these labels should be propagated to processed images and catalog data.}

\textbf{Region Labels -- } 
Based on Right Acension and Declination.
To be applied to data by Rubin Data Management, e.g., using a look-up table.
Region labels would include, e.g., WFD low-dust, Galactic Plane, Galactic 
Bulge, SCP, NES, or Virgo; DDF fieldname; mini-survey region name.
Region labels should be propagated to processed images (visit, 
difference, and deep coadds), catalog sources, and catalog static-sky objects.

\textbf{Observing Mode Labels -- } 
Based on the scheduler mode at the time of the observation
(i.e., survey cadence or strategy).
To be applied by the Survey Scheduler team via the observations scheduler.
Observing mode labels might include, e.g., WFD, DDF, TOO, mini-survey name, or other 
options like engineering, commissioning, and director's discretionary.
Observing mode  labels should be propagated to single-visit processed and 
difference images and catalog sources.
It would not make sense to propagate observing mode labels to deep coadds or 
object tables, as these data products could be a mix of modes.

@@

\item To enable provenance when Special Programs data in included in regular
processing and the Main Survey's data products (e.g., if used to improve the
all-sky coadd).

@@

\textbf{Necessary -- } When the primary science goal for a Special Program 
cannot be met by including the data in regular processing (e.g., Prompt processing),
or where doing so would compromise the Main Survey data products (e.g., introduce non-uniformity).

@@

The main motivation for this recommendation is that all time-domain and 
moving-object science goals (two of the four science pillars for the LSST) are
enhanced by any and all additional observations, even if they are not
optimized for the science (as the WFD area's cadence is being optimized).

@@

Users will need to be able to query for data that restricts by sky region and 
observign mode label. 
This can be accomplished by including those labels in all image and catalog 
metadata as described in Section~\ref{ssec:sci_labels}, as the TAP service 
and butler already provide the mechanism for user-specified queries.

Users will need to be able to discover Special Programs data products when 
browsing data.
This could be accomplished with, e.g., a toggle to overlay region boundaries
for Special Programs when browsing all-sky coadds built from Main Survey data
in the Portal Aspect.
